---
title: "Practical Machine Learning Prediction Assignment"
author: "Sunil V Nair"
date: "15 October 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

### Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set.

### Loading the Dataset

```{r libs}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(gbm)
library(plyr)
```

```{r readLoad}
# Download the training data

myfile<-'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv?accessType=DOWNLOAD'

# Load the training dataset
data_training <- read.csv(myfile, na.strings=c("NA","#DIV/0!",""))

# Download the testing data


myfile<-'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv?accessType=DOWNLOAD'

# Load the testing dataset
data_testing <- read.csv(myfile, na.strings=c("NA","#DIV/0!",""))
```

###Cleaning the Data
In this section, we will remove all columns that contains NA and remove features that are not in the testing dataset. The features containing NA are the variance, mean and standard devition (SD) within each window for each feature. Since the testing dataset has no time-dependence, these values are useless and can be disregarded. We will also remove the first 7 features since they are related to the time-series or are not numeric.
```{r clnData}
features <- names(data_testing[,colSums(is.na(data_testing)) == 0])[8:59]

# Only use features used in testing cases.
data_training <- data_training[,c(features,"classe")]
data_testing <- data_testing[,c(features,"problem_id")]

dim(data_training); 
dim(data_testing);
```

###Partitioning the Dataset
Following the recommendation in the course Practical Machine Learning, we will split our data into a training data set (60% of the total cases) and a testing data set (40% of the total cases; the latter should not be confused with the data in the pml-testing.csv file). This will allow us to estimate the out of sample error of our predictor.
```{r partData}
set.seed(12345)

inTrain <- createDataPartition(data_training$classe, p=0.6, list=FALSE)
training <- data_training[inTrain,]
testing <- data_training[-inTrain,]

dim(training); 
dim(testing);
```

###Decision Tree Model
Using Decision Tree, we shouldn't expect the accuracy to be high. In fact, anything around 80% would be acceptable.
```{r treem}
set.seed(12345)
dttree <- rpart(classe ~ ., data = training, method="class", control = rpart.control(method = "cv", number = 10))
fancyRpartPlot(dttree)
```

###Predicting using Decision Tree Model
```{r predTree}
set.seed(12345)

prediction <- predict(dttree, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```

###Random Forest Model
Using random forest, the out of sample error should be small. The error will be estimated using the 40% testing sample. We should expect an error estimate of < 3%.
```{r rfm}
set.seed(12345)

dtRfm <- randomForest(classe ~ ., data = training, method = "rf", importance = T, trControl = trainControl(method = "cv", classProbs=TRUE,savePredictions=TRUE,allowParallel=TRUE, number = 10))

plot(dtRfm)
```

###Boosting Model
```{r boostm}
boostM <- train(classe ~ ., method = "gbm", data = training, verbose = F, trControl = trainControl(method = "cv", number = 10))

boostM
```

```{r Pltboostm}
plot(boostM)
```

###Predicting using Random Forest Model
```{r predRfm}
prediction <- predict(dtRfm, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```

Accuracy :- 99.3% 

###Predicting using Boosting Model
```{r predBoostm}
prediction <- predict(boostM, testing)
confusionMatrix(prediction, testing$classe)
```

###Predicting with the Testing Data (pml-testing.csv)
####Decision Tree Prediction
```{r dft}
predictionDT <- predict(dttree, data_testing)
predictionDT
```

####Random Forest Prediction
```{r rfp}
predictionRF <- predict(dtRfm, data_testing)
predictionRF
```

####Boosting Prediction
```{r prfp}
predictionBoost <- predict(boostM, data_testing)
predictionBoost
```



###Conclusion &  file Submission
As can be seen from the confusion matrix the Random Forest model is very accurate, about 99%. Because of that we could expect nearly all of the submitted test cases to be correct. It turned out they were all correct.

Prepare the submission.
```{r pwriteFile}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("pwriteFile_problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictionRF)

predictionRF
```

